\chapter{Creating an Avatar}

This section is for building a 3D avatar using a specific group of software tools.

\section{Signing Avatar Development Technologies}


Unity is a cross-platform graphics engine developed by Unity Technologies, specializing in creating 2D and 3D video games and audio and animation rendering. This engine offers networking tools for multiplayer experiences, NavMesh navigation tools for Artificial Intelligence, Virtual Reality support, and real-time interactive content creation. Based on the C\# programming language, Unity brings graphics to life in the game. Its robust documentation and active user community are remarkable, providing resources and support for developers. For Windows, Mac OS, and Linux, Unity has been widely used in various games thanks to its versatility in adapting to different systems, creating 2D and 3D scenes with real-time animations and high-quality graphics performance, and providing an immersive experience. Unity offers both a free and a paid version with more advanced features.\\

MakeHuman is a user-friendly tool crafted to integrate the process of building virtual human characters through a Graphical User Interface (GUI). MakeHuman aims to simplify and enhance the creation of lifelike virtual humans, providing an efficient approach to this particular branch of digital design. \\

Ready Player Me accommodates full and half-body avatars, specifically optimized to enhance real-time game performance and uphold quality standards. This adaptable avatar system is crafted for games, VR/AR experiences, and applications. Furthermore, it seamlessly aligns with various technologies and game engines, including Unity, Unreal Engine, React, native mobile solutions, and API integration with stacks supporting REST and PostMessage protocols. Incorporating the Ready Player Me Avatar creator into your application or game is user-friendly. With many customization options, users can personalize avatars to reflect their identities genuinely. These avatars are supplied fully skinned, rigged, and primed for animation, ensuring a smooth integration into the preferred development environment. For a streamlined integration experience, developers can leverage the provided SDKs (Unity, Unreal, React) or engage directly with the APIs and/or iFrame Avatar Creator integration. The Ready Player Me avatar system functions as an all-encompassing solution for creating, managing, integrating, and customizing player avatars across diverse platforms, encompassing web, desktop, mobile, VR, and others supported by Unity and Unreal Engine. \\


Blender stands out as a comprehensive, free, and open-source 3D creation suite that spans the entire spectrum of the 3D pipeline. This includes modeling, rigging, animation, simulation, rendering, compositing, motion tracking, video editing, and even game creation. Operating as a cross-platform application, Blender seamlessly functions on Linux, Windows, and Macintosh computers. Its interface, powered by OpenGL, ensures a consistent and user-friendly experience.

With Blender, users can unleash their creativity to generate diverse 3D visualizations encompassing still images, 3D animations, and VFX shots. The software is built on a high-quality 3D architecture, facilitating a speedy and efficient workflow for content creation.

What sets Blender apart is not only its robust features but also its vibrant community support. Users benefit from an active community that contributes to ongoing improvements, making Blender a powerful and accessible tool for 3D enthusiasts across the globe. 

Blender is considered as a tool for this research as it
provides several features as mentioned before that will support in the development of an embodied conversational agent. Blender provides its avatar rendering engine, which
provides capabilities for real-time processing. 